---
title: "Africa RISING Locations"
author: "Zambia"
date: "DRAFT 7/12/2016"
output: html_document
---

```{r, echo=F, cache=T, warning=F}

library(rgdal)
library(leaflet)
library(raster)
library(rhandsontable)

# Load data
load("./data/ARPoints.RData")

```

Currently showing `r nrow(ar[ar$country_id=="ZMB",])` camps and `r nrow(zmb)` villages. Village points are colored by camp (i.e. each camp is a different color). In bright red on the map and in the table below are villages with -- possibly -- bad coordinates (please check and report back).


```{r, echo=F, cache=T}

# Generate leaflet map
leaflet(width="100%", height=640) %>%
  setView(30,-13, 6) %>%
  
  addTiles(
    urlTemplate="https://2.aerial.maps.api.here.com/maptile/2.1/maptile/newest/hybrid.day/{z}/{x}/{y}/256/png8?app_id=ViVS62MmsdTJSzwXMlHb&app_code=bsODmSCzkVqEMR8TnGPHdw",
    attribution="&copy; HERE maps, 2014 DigitalGlobe.",
    group="HERE Hybrid Day") %>%
  addTiles(
    urlTemplate="http://a.tile.opentopomap.org/{z}/{x}/{y}.png",
    attribution="&copy; OpenStreetMap, SRTM | &copy; OpenTopoMap (CC-BY-SA)",
    group="OpenTopoMap") %>%
  addTiles(
    urlTemplate="http://server.arcgisonline.com/ArcGIS/rest/services/Specialty/DeLorme_World_Base_Map/MapServer/tile/{z}/{y}/{x}",
    attribution="Tiles &copy; Esri &mdash; Copyright: &copy;2012 DeLorme",
    group="ESRI DeLorme") %>%
  
  # Add rasters
  addRasterImage(alt, opacity=.8, group="Altitude", colors=pal.elev) %>%
  addRasterImage(strat.zmb, opacity=.8, group="Stratification", colors="RdYlBu") %>%
  
  # Add villages
  addCircleMarkers(data=zmb,
    fillColor=~colorFactor("Spectral", NULL)(camp), 
    group="Villages",
    label=~paste(village_id, paste("Village:", village), paste("Camp:", camp), sep=" | "), 
    popup=~paste(
      paste("Region:", region), 
      paste("District:", district), 
      paste("Camp:", camp),
      paste("Village:", village), sep="<br>"),
    color="white", weight=2, fillOpacity=1, radius=7,
    #clusterOptions=markerClusterOptions(),
    options=markerOptions(riseOnHover=T)) %>%
  
  # Add bad villages
  addCircleMarkers(
    data=zmb[zmb$village %in% bad,],
    group="Villages (problems?)",
    label=~paste(village_id, paste("Village:", village), paste("Camp:", camp), sep=" | "),
    fillColor="red", color="white", weight=2, fillOpacity=1, radius=8, dashArray="5, 5",
    options=markerOptions(riseOnHover=T)) %>% 
  
  # Add camps
  addMarkers(
    data=ar[ar$country_id=="ZMB",],
    group="Camps",
    label=~paste("Camp:", village),
    popup=~paste(
      paste("<strong>Region</strong>:", region), 
      paste("<strong>District</strong>:", district), 
      paste("<strong>Lead partner</strong>:", partner), 
      paste("<strong>Partners</strong>:", partners),
      paste("<strong>Activities</strong>:", site_desc), sep="<br>"),    
    options=markerOptions(riseOnHover=T)) %>%  
  
  # Add non-Zambia locations
  addCircleMarkers(
    data=ar[ar$country_id!="ZMB",],
    fillColor="grey40", 
    group="Other AR Locations",
    label=~paste(village_id, 
      paste("Village:", village), paste("District:", district), sep=" | "), 
    popup=~paste(
      paste("Region:", region), 
      paste("District:", district),
      paste("Village:", village), sep="<br>"),
    color="white", weight=1, fillOpacity=1, radius=5,
    #clusterOptions=markerClusterOptions(),
    options=markerOptions(riseOnHover=T)) %>%    
  
  # # Legend
  # addLegend(title="Zambia",
  #   opacity=1,
  #   position="bottomleft", 
  #   colors=c("green", "red"),
  #   labels=c("village (GPS seem OK)", "village (with bad GPS)"))  %>%
  
  # addLegend(title="Camps",
  #   opacity=1, pal="Spectral",
  #   position="bottomleft", 
#   values=aslevels(zmb$site_desc),
#   labels=levels(zmb$site_desc))  %>%  

  # Add layer controls
  hideGroup(c("OpenTopoMap", "HERE Hybrid Day", "Altitude", "Stratification", "Villages (problems?)")) %>%
  addLayersControl(
    baseGroups=c("OpenTopoMap", "HERE Hybrid Day", "ESRI DeLorme"),
    overlayGroups=c("Villages", "Villages (problems?)", "Camps", "Other AR Locations", 
      "Altitude", "Stratification"),
    options=layersControlOptions(collapsed=F)) %>%
  addMiniMap(zoomLevelFixed=4, autoToggleDisplay=F) %>%
  addScaleBar("bottomright")


```


## Villages with *Problem* Coordinates

These `r nrow(zmb[zmb$village %in% bad,])` villages are either located outside Zambia, or are quite remote from the rest of the camp, else they could be duplicates.

```{r, echo=F}

rhandsontable(zmb@data[zmb$village %in% bad,], 
  rowHeaders=NULL, readOnly=T, width="100%", 
  contextMenu=F, highlightRow=T, columnSorting=T)


```


<small>Source on [GitHub](https://github.com/harvestchoice/hc-shiny/tree/master/ar)</small>.

