---
title: "HarvestChoice - CELL5M Data Validity Tests"
author: "Bacou, Melanie"
date: "7/31/2015"
runtime: shiny
output: 
html_document: 
toc: true
fig_height: 6
fig_width: 6
---

A sample of data validation tests for HarvestChoice 5-arc minute spatial layers for sub-Sharan Africa (documented at http://harvestchoice.org/data). These are meant to catch more obvious processing errors, not so much to proof the quality of the original sources. These tests are re-ran every time this page is loaded.

```{r, results='hide'}
setwd("/home/projects/cell5m")

# Load libraries
library(data.table)
library(raster)
library(hcapi3)
library(tmap)
library(stringr)
library(leaflet)
library(shiny)

```

## Administrative Classification

Verify countries, admin-1, and admin-2 classification across SSA grid. For now this classification should be consistent with FAO GAUL 2008 (2009 eds), aside from Kenya for which we use the 2013 47-county classification. We can load GAUL at level-2 and compare with CELL5M.

```{r, results='hide'}
# Load GAUL admin-2
g2 <- readRDS("/home/projects/cell5m/rdb/g2_2008v09.web.rds")
names(g2)
g2.dt <- data.table(g2@data)
g2.dt[, rn := row.names(g2)]

```

```{r}
# Check wich version of Kenya this is
g2.dt[ADM0_NAME=="Kenya", .N]
g2.dt[ADM0_NAME=="Kenya", .N, keyby=ADM1_NAME]

# Compare against Kenya boundaries found in CELL5M
# Get the needed variables from CELL5M
dt <- getLayer(c("ADM0_CODE", "ADM1_CODE_ALT", "ADM2_CODE_ALT", "AREA_TOTAL"))
names(dt)
dt[ADM0_NAME=="Kenya", .N]
ken <- dt[ADM0_NAME=="Kenya", list(ADM2_NAME=unique(ADM2_NAME_ALT)), keyby=ADM1_NAME_ALT]
ken
nrow(ken[ADM1_NAME_ALT!="buffer gridcell"])
```

47 counties are shown for Kenya as opposed to 86 in the official GAUL 2008 (2009) release. This is normal. We can verify all other country classification similarly.

### Look for Missing Admin Units

```{r}
# For every country summarize the number of districts across provinces in GAUL and in CELL5M, and compare. We remove all "buffer gridcells" from the CELL5M summary.
g2.admin <- g2.dt[, list(ADM0_NAME, ADM1_NAME, ADM2_NAME), 
  keyby=list(ADM0_CODE, ADM1_CODE, ADM2_CODE)]
dt.admin <- dt[ADM1_NAME_ALT!="buffer gridcell", .N, 
  keyby=list(ADM0_CODE, ADM1_CODE_ALT, ADM2_CODE_ALT)]

# Combine and look for any unmatched districts
g2.admin$match <- dt.admin[g2.admin][, N]
summary(g2.admin$match)
unmatched <- g2.admin[is.na(match) & ADM0_NAME != "Kenya", .N, 
  by=list(ADM0_NAME, ADM1_CODE, ADM1_NAME, ADM2_NAME, ADM2_CODE)]

unmatched[, unique(ADM0_NAME)]

```

`r nrow(unmatched)` GAUL districts were not found in CELL5M across `r unmatched[, unique(ADM0_NAME)]` countries. We can map them.

```{r, fig.width=3, fig.height=3, warning=F}
# Plot missing units in each country
p <- lapply(unique(unmatched$ADM0_NAME), function(i) {
  u <- unmatched[ADM0_NAME==i, ADM2_CODE]
  cntr <- g2[g2$ADM0_NAME==i,]
  cntr$col <- ifelse(cntr$ADM2_CODE %in% u, T, F)
  
  # Zoom in towards the missing units (they can be very small)
  b <- extent(cntr[cntr$ADM2_CODE %in% u,])+c(-3,3,-3,3)
  
  tm_shape(cntr, bbox=bbox(b)) +
    tm_polygons(col="col", border.col="white", palette=c("grey90", "red"),
      labels=c("matched", "unmatched"), title=i) +
    tm_layout(title.size=.8,
      title=str_wrap(
        paste(unmatched[ADM2_CODE %in% u, ADM2_NAME], collapse=", "), width=30))
})

for (i in p) print(i)

```


In all instances these seem to correspond to city limits that were too small to intersect with CELL5M grid. We could correct those manually to make sure they are accounted for.

Because we keyed and matched GAUL and CELL5M on a combination of `ADM0_CODE`, `ADM1_CODE` and `ADM2_CODE` the unmatched table would also return districts that might be classified within the wrong provinces and/or countries. No such record was returned though, so seems correct (e.g. there is no case of districts in Lesotho showing as belonging to South Africa).

Another way to check visually is to print all districts as a color-coded raster by country. We can display these layers in a Leaflet widget.

```{r, results="hide"}
# Raster needs all character variables categorized into factors first
dt[, `:=`(
  ADM0_NAME=factor(ADM0_NAME),
  ADM1_NAME_ALT=factor(ADM1_NAME_ALT),
  ADM2_NAME_ALT=factor(ADM2_NAME_ALT),
  ISO3=factor(ISO3))]
```

```{r, warning=F}
# Convert CELL5M variables to a raster brick (multi-layer raster object)
r <- SpatialPixelsDataFrame(dt[, list(X,Y)], data.frame(dt),
  proj4string=CRS("+init=epsg:4326"),  tolerance=0.0036002)
r <- brick(r)
names(r)

# Reproject layers to epsg:3857 to speed up leaflet
r <- projectRaster(r, crs=CRS("+init=epsg:3857"))
g2 <- spTransform(g2, CRS("+init=epsg:3857"))

```


```{r}
# Create leaflet widget
m <- renderLeaflet({
  # Crop extent to selected country
  d <- crop(r, g2[g2$ADM0_NAME==input$selectADM0,])
  p <- dt[ADM0_NAME==input$selectADM0, list(X=mean(X), Y=mean(Y)),
    by=list(ADM0_CODE, ADM1_CODE_ALT, ADM2_NAME_ALT)]
  
  # Update map
  leaflet() %>%
    setView(mean(p$X), mean(p$Y), 4) %>%
    addTiles(urlTemplate="http://{s}.tiles.mapbox.com/v3/jcheng.map-5ebohr46/{z}/{x}/{y}.png") %>%
    addRasterImage(raster(d, "ADM0_NAME"), opacity=.6, project=F, group="Country") %>%
    addRasterImage(raster(d, "ADM1_NAME_ALT"), opacity=.6, project=F, group="Provinces") %>%
    addRasterImage(raster(d, "ADM2_NAME_ALT"), opacity=.6, project=F, group="Districts") %>%
    addMarkers(data=p, lng=~X, lat=~Y, popup=~ADM2_NAME_ALT, 
      clusterOptions=markerClusterOptions(),
      options=markerOptions(riseOnHover=T)) %>%
    hideGroup(c("Provinces", "Districts")) %>%
    addLayersControl(overlayGroups=c("Country", "Provinces", "Districts"),
      options=layersControlOptions(collapsed=F))
})

# Render user input
selectInput("selectADM0", "Select a Country", choices=levels(dt$ADM0_NAME), selected="Ghana")

```

```{r, echo=F}
m
```


## Published Layers and Metadata Completeness

Another batch of quality tests is to look for missing metadata in the Variable Inventory table `vi`. This tables is also part of the HC Data API and was loaded above with `library(hcapi3)`.

Look for possibly misformed raster legends.


## Analytical Features

### Domain Summaries

### Top/bottom Ranking






