---
title: "HarvestChoice - CELL5M Data Validation"
author: "Bacou, Melanie"
date: "8/1/2015"
runtime: shiny
output: 
html_document:
fig_height: 6
fig_width: 6
---

This document is part of a series of automated tests to validate [HarvestChoice](http://harvestchoice.org/data) 5-arc-minute spatial rasters for sub-Saharan Africa.

**Contents**

- [Administrative Classification](index.Rmd)
- [Metadata Summary](viValidate.Rmd)
- [Variable Inventory](viValidate.Rmd#variable-inventory)
- [Domain Summary](#domain-summary)
- [Top/Bottom Ranking](#top-bottom-ranking)
- [Market Analysis](#market-analysis)
- [Homologue Tool](#homologue-tool)


```{r, eval=F}

# Load common libraries
library(data.table)
library(hcapi3)
```

```{r}

library(rhandsontable) # Just to render fancy tables on this page

```

## Basic Data Consistency Tests

```{r, echo=F}

vars <- vi[published==T & type=="continuous", list(cat3=unique(cat3)), by=list(cat1, cat2)]
vars <- split(vars, vars$cat1)
vars <- lapply(vars, function(x) split(x, x$cat2))
vars <- lapply(vars, function(x) lapply(x, function(i) i$cat3))
selectInput("selectCat", "Select a variable", vars, selected="Precipitation")

```

```{r}

# Choose to quickly visualize subsets of variables (across sub-categories, limited to 6)
renderPlot(height=900, {
  p <- vi[order(sortOrder)][cat3==input$selectCat, varCode]
  if (length(p) > 6) p <- sample(p, 6)
  if (length(p) > 1) par(mfrow=c((length(p)%/%2)+(length(p)%%2), 2))
  for (i in p) hcapi(i, format="stats")
})

# Rank these same variable(s) across countries
renderRHandsontable({
  p <- vi[order(sortOrder)][cat3==input$selectCat, varCode]
  df2 <- hcapi(p, by="ADM0_NAME")
  df2 <- df2[, lapply(.SD, function(x) df2[order(x, decreasing=T)[1:30], ADM0_NAME]), .SDcols=-1]
  setnames(df2, vi[names(df2)][, varLabel])
  rhandsontable(df2, height=420, width="100%", stretchH="all") %>%
    hot_cols(colWidths=100)
})

```



## Test Analytical Methods

We run a series of pre-defined tests across HC API analytical mehods and show the results here interactively.

### Domain Summary

#### Simple Summary across Zones

All rasters are summarizable across zones (e.g. across farming systems, agro-ecological zones, administrative boundaries, market access zones), or across any geographic domain defined by a categorized raster (by default a continuous raster will be classified along the legend breaks found in the metadata). In addition users nay pass their own breaks, e.g. using `hcapi("bmi", by=list(whea_r_h=c(100, 200, 400, 800, 1600, 3200, 6400)))`.

```{r}

# For example, show default legend breaks for rainfed maize harvested area
renderRHandsontable({
  df <- vi["maiz_r_h"][, lapply(.SD, function(x) unlist(strsplit(x, "|", fixed=T))), 
    .SDcols=c("classLabels", "classBreaks", "classColors")]  
  rhandsontable(df, width=500, stretchH="all")
})

```

```{r, echo=F}

# Add user select
selectInput("selectISO", "Select a country", iso)

```

```{r}

# Summarize 5 random rasters across zones of low to high maize harvested area using default breaks.
vars1 <- sample(vi[published==T & varCode != "maiz_r_h", varCode], 5)

renderRHandsontable({
  df <- hcapi(vars1, iso3=input$selectISO, by="maiz_r_h")
  setnames(df, vi[names(df)][, varLabel])    
  rhandsontable(df, height=240, width="100%", stretchH="all") %>%
    hot_cols(colWidths=100, type="numeric", format="0,0.0#", renderer=convertNA())
})

# Same using custom breaks on cassava zones
renderRHandsontable({
  df <- hcapi(vars1, iso3=input$selectISO,
    by=list(cass_h=c(100, 200, 400, 800, 1600, 3200, 6400)))
  setnames(df, vi[names(df)][, varLabel])    
  rhandsontable(df, height=200, width="100%", stretchH="all") %>%
    hot_cols(colWidths=100, type="numeric", format="0,0.0#", renderer=convertNA())
})

```

```{r}

# Summarize another 4 random rasters across districts for the same country
vars2 <- sample(vi[published==T, varCode], 4)

renderRHandsontable({
  df <- hcapi(vars2, iso3=input$selectISO, by="ADM2_NAME_ALT")
  setnames(df, vi[names(df)][, varLabel])
  rhandsontable(df, height=400, width="100%", fixedColumnsLeft=1, stretchH="all") %>%
    hot_cols(colWidths=100, type="numeric", format="0,0.0#", renderer=convertNA())
})

```




#### Summary across Custom Geographies

### Top/bottom Ranking

### Market Analysis

### Homologue Tool



